---
title: "Data Transformation in R"
author: "Loris Moraiti"
date: "2023-10-25"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This document explains how to use the `transform_data` function in R to perform two different data transformations on a dataset.

```{r function, echo=TRUE}
# Define a function to perform different data transfromations
transform_data <- function(data, transformation_type = "standardize") {
  numeric_columns <- sapply(data, is.numeric)
  if (transformation_type == "standardize") {
    data[, numeric_columns] <- scale(data[, numeric_columns])
  } else if (transformation_type == "log_transform") {
    data[, numeric_columns] <- log(data[, numeric_columns] + 1) # Adding 1 to avoid log(0)
  }
  
  return(data)
}
```

# Function Overview

The `transform_data` function has the following functionality:

- It takes two arguments:
  1. `data`: The dataset you want to transform.
  2. `transformation_type`: A string that determines the type of transformation to perform. It can be one of the following:
    - `"standardize"`: Standardizes the numeric columns to have a mean of 0 and a standard deviation of 1.
    - `"log_transform"`: Applies a natural logarithm transformation to the numeric columns, with a small constant added to avoid taking the logarithm of zero.

# Usage

To use the `transform_data` function, follow these steps:

1. Load your dataset. For example, you can use the built-in mtcars dataset.

```{r cars, echo=FALSE}
# Load the mtcars dataset
data(mtcars)
```

2. Call the 'transform_data' function with your dataset and specify the transformation type. Depending on the 'transformation_type' argument, the function outputs different results. These are displayed below.

## transformation_type = "standardize"

```{r cars standardize, echo=FALSE}
# Standardize the data
standardized_mtcars <- transform_data(mtcars, transformation_type = "standardize")
head(standardized_mtcars)
```

## transformation_type = "log_transform"

```{r cars log, echo=FALSE}
# Log-transform the data
log_transformed_mtcars <- transform_data(mtcars, transformation_type = "log_transform")
head(log_transformed_mtcars)
```

## Conclusion

This R Markdown document demonstrates how to use the transform_data function and highlights the different outputs depending on the value of the transformation_type argument. You can use this function to tailor data transformations to your specific analysis needs.

## Appendix: All code in this assignment

```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE} 
# this chunk generates the complete code appendix. 
# eval=FALSE tells R not to run (``evaluate'') the code here (it was already run before).
```